---
title: "Assignment NYC Flight Data"
author: "Filza Mazahir"
date: "October 11, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYC Flight Data

Your job is to create a "rectangular" table useful for modeling from flight data. In the folder `02-fundamentals/data` there are four CSV files: `flights.csv`, `airports.csv`, `planes.csv` and `weather.csv`.  Put code in each of the sections provided.

## 1. Read Data

Using the `readr` package read the `flights` data.

```{r}
library(readr)
flights <- read_csv("data/flights.csv")

```

## 2. Numeric Variables  

Plot a histogram of arrival delays and departure delays 

```{r}
hist(flights$arr_delay, 
     main="Arrival Delays", 
     xlab="Arrival delay time in minutes", 
     border="green", 
     col="blue")

hist(flights$dep_delay, 
     main="Departure Delays", 
     xlab="Departure delay time in minutes", 
     border="blue", 
     col="green")


```

## 3. Scatterplot

Plot a scatterplot of date vs arrival delay 

```{r}
flights$date <- as.Date( paste(flights$month , flights$day , sep = "." )  , format = "%m.%d" )

plot(flights$date, flights$arr_delay,
     xlab="Date",
     ylab="Arrival Delay")
```



## 4. Create tables for Categorical Variables

Create a `table` of occurences of origin airports and carrier.

```{r}
library(plyr)
occurences <- count(flights, c('origin', 'carrier'))
print(occurences)

```


## 5. Join Data

**Read** in the other data sets. Use methods to join the data together to create a single table containing one record per row.  (HINT: see `?data.table` or `?dplyr::join`)

```{r}
library(dplyr)
airports <- read_csv("data/airports.csv")
planes <- read_csv("data/planes.csv")
weather <- read_csv("data/weather.csv")

flights_planes <- flights %>% left_join(planes, by = "tailnum") 

flights_planes_weather <- flights_planes %>% left_join(weather, by= c("origin", "time_hour"))

joined_table <- flights_planes_weather %>% left_join(airports, c("origin" = "faa"))
print(joined_table)
```



