---
title: "NYCFlights: Arrival Delay Logictic Model"
author: "Filza Mazahir"
date: "October 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(readr)
library(dplyr)
library(psych)
```


## Logsitic and Inverse Logistic Transformation 

- Write an R function for the logistic function. The function should accept a `numeric` vector with values `[-Inf,Inf]` and produce a numeric vector in the the range `[0,1]`.

- Plot the logistic function from  `[-10,10]`

- Write a R function for the inverse logistic function. The function should accept a `numeric` vector with values `[0,1]` and prodcuce a numeric vector in the range `[-Inf,Inf]`

- Plot the Inverse Logistic function from `[0,1]`


**Hint:** For plotting curves see `?graphics::curve` or `?ggplot2::stat_function`

```{r "Logistic and Inverse Logistic" }
logisticfunction <- function(x){1/(1 + exp(-x))}
ggplot(data.frame(x = c(-10, 10)), aes(x)) + stat_function(fun = logisticfunction)

inverselogisticfunction <- function(x){log(x/(1-x))}
ggplot(data.frame(x = c(0, 1)), aes(x)) + stat_function(fun = inverselogisticfunction)

```


# NYCFlights Model

Using the rectangular data that you created from assignment and following the *Predicting Medical Expenses* example from the text (*MLwR*), create a model for arr_delay >= 15 minutes. Follow *MLwR* structure   for building a model. Describe/Explain each of the steps and show all work in codeblocks below.

KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file.   

## Step 1: Collect Data 

```{r "Step 1: Collect Data" }
flights <- read_csv("data/flights.csv")
airports <- read_csv("data/airports.csv")
planes <- read_csv("data/planes.csv")
weather <- read_csv("data/weather.csv")


yx <- 
  flights %>% 
  left_join(weather, by=c("origin","year","month","day","hour") ) %>%
  left_join(planes, by='tailnum', suffixes=c('','.pl') ) %>% 
  left_join(airports, by=c('origin'='faa') ) %>% 
  left_join(airports, by=c('dest'='faa') )  



yx$arr_delay_binary<-ifelse(yx$arr_delay<15,0,1)

write.csv(yx, file = "yx.csv")
  
```

## Step 2: Explore and Prep The Data


One of the things not done in the MLwR text is a pairwise comparison between the response and each of the predictors. Make sure to do this; this is often very illustrative of the relationship between that predictor and the response. This can be done with `pairs` or `psych::panel.pairs`


```{r "Step 2: Explore and Prep The Data" }

##print(pairs.panels(yx[c('month', 'dep_delay', 'carrier', 'air_time', 'distance' , 'year.y', 'type', 'engine', 'wind_dir', 'wind_speed', 'wind_gust', 'precip', 'pressure', 'visib', 'lat.x', 'lon.x', 'lat.y', 'lon.y', 'arr_delay_binary')]))



```

## Step 3:  Train The Model

```{r "Step 3: Train The Model" }

```

## Step 4: Evaluate Performance

Think about some of the tools that you have for evaluating performance.  Choose one and articulate why you have chosen it.

```{r "Step 4: Evaluate Performance" }

```

 
## Step 5: Improve Performance 

Show some steps for improving model performance.

```{r "Step 5: Improve Performance" }

```


# Question:

Is this a good model?  (Write your answer here.)


# PART B:

Your model should be good at explaining tardiness. Now, assume that your job is to predict arrival delays a month in advance. You can no longer use all the features in your model. Retrain your model using only features that will be *known* only a month in advance of the departure time.  Show all steps as above.

